% Created 2024-12-07 Sat 18:34
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{parskip,darkmode}
\enabledarkmode
\author{Arnav Gupta}
\date{\today}
\title{Microservices}
\hypersetup{
 pdfauthor={Arnav Gupta},
 pdftitle={Microservices},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.4 (Org mode 9.7.11)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\section{Monolithic Design}
\label{sec:org62f80eb}
One big application.
Each instance can handle any request on its own (other than help from DB).

When developer is testing new code, they can run one instance locally.

Monolithic design is easy to build, deploy, test, coordinate, and share data.

Monolithic disadvantages:
\begin{itemize}
\item creates bottlenecks during development
\begin{itemize}
\item many devs working on one codebase
\end{itemize}
\item big codebase can lead to messy, fragile code
\item whole app must be redeployed for small new functionality
\item must choose one programming language, build system, runtime env
\end{itemize}
\section{Microservices}
\label{sec:orgb39c896}
In \textbf{Service Oriented Architecture}, many smaller services work together.

Responsibilities split among specialized services.

To outside world, looks like one application.

Each \textbf{microservice} is a black box to the rest of the system, and handles requests from the network
(implementation hidden).
Language-independent network-level API needed to specify format of requests and responses.

Microservice disadvantages:
\begin{itemize}
\item more difficult to trace request handling code for debugging, since it spans many apps
\begin{itemize}
\item but easier to test smaller services
\end{itemize}
\item ACID transactions more difficult to support
\begin{itemize}
\item monolithic design can manage a single DB connection with a transaction that can be rolled back
\end{itemize}
\item devs get isolated in sub-projects, blocking collaboration and innovation
\end{itemize}

Microservices isolate codebases with clear network API boundares, allowing parallel work.

When starting a new project:
\begin{itemize}
\item organize system into services and DBs
\item agree on network-level API for each service
\item assign engineers to each service and build independently
\end{itemize}

If application interacts with service under construction, can build \textbf{mock} of service (returns
hard-coded data for testing).

Traditional web-app generates HTML on server.
For JS single-page app, JS app runs in browser and makes REST requests and generates HTML.

Service API (REST, GraphQL) essential for supporting mobile and desktop apps with cloud-based data
and services.

Single API can handle all client types.
\end{document}
