% Created 2024-12-10 Tue 15:00
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{parskip,darkmode}
\enabledarkmode
\author{Arnav Gupta}
\date{\today}
\title{Quality Attributes}
\hypersetup{
 pdfauthor={Arnav Gupta},
 pdftitle={Quality Attributes},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.4 (Org mode 9.7.11)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\section{Quality Attributes}
\label{sec:org2d61fed}
Includes \uline{reliability}, \uline{availability}, \uline{portability}, \uline{scalability},
and \uline{performance}.

\textbf{Information Capture and Dissemination Environment (ICDE)}: software system for
providing intelligent assistance to analysts and researchers

ICDE automatically captures and stores data of actions performed by a user.
Data can later by used by 3rd parties to offer intelligent help.

ICDE business goals:
\begin{itemize}
\item encourage 3rd party tool developers
\begin{itemize}
\item \uline{integration}: simple and reliable programmatic access to data store for
3rd party tools
\item \uline{portability}: heterogeneous platform support for running 3rd party
tools
\item \uline{reliability}: allow 3rd party tools to communicate with ICDE users from a
remote machine
\end{itemize}
\item promote ICDE concept to users
\begin{itemize}
\item \uline{scalability}: scale data collection and data store components to support
up to 150 users at a single site
\item \uline{scalability}: low-cost deployment for each ICDE user workstation
\end{itemize}
\end{itemize}

Quality attributes must be made precise/measurable for a given system design.
\subsection{Performance}
\label{sec:orgb90df7a}
Different ways to measure performance include: throughput, response time,
deadlines

Enterprise applications often have strict performance requirements.
\subsubsection{Throughput}
\label{sec:orgbe0bbbb}
Measure of amount of work an application must perform in unit time.

Required throughput can be average or peak.

Many systems have low average but high peak throughput requirements.
\subsubsection{Response Time}
\label{sec:org7590cb9}
Measure of latency an application exhibits in processing a request.

Usually measured in milliseconds, and often an important metric for users.

Required response time can be guaranteed or average.
\subsubsection{Deadlines}
\label{sec:org80fb85a}
Something must be completed before some specified time.
Most often associated with batch jobs in IT systems.
\subsubsection{Issues}
\label{sec:org41bc076}
Overheads of trapping user events must be imperceptible to ICDE users
(response time).
Solution is to decouple user event capture from storage using a queue.
\subsection{Scalability}
\label{sec:org4cfdd03}
Main scalability issues are request load, connections, data size, and deployments
\subsubsection{Request Load}
\label{sec:org41de869}
How an application behaves when simultaneous load grows.

Ideal solution, without additional hardware capacity is: as the load increases, throughput remain
constant and response time per request increases only linearly.

Can scale by adding more hardware: scale-out \(\to\) horizontal or scale-up \(\to\) vertical

In reality, applications will exhibit a decrease in throughput and a subsequent exponential increase
in response time:
\begin{itemize}
\item increased load causes increased \uline{resource contention} (CPU, network, memory)
\item each request consumes some additional resource (buffer space, locks, etc) in the application,
and eventually these are exhausted
\end{itemize}
\subsubsection{Connections}
\label{sec:org9f9df95}
When the number of simultaneous connections to an application increases.
Each connections consumes some resources and this can exceed the maximum number of connections.
\subsubsection{Data Size}
\label{sec:orgdf516a3}
How an application behaves as the data it processes increases in size.
\subsubsection{Deployment}
\label{sec:orgc47caaa}
How effort to install/deploy an application increases as installation base grows, by installing new
users or new servers.

Solutions typically revolve around automatic download/installation.
\subsubsection{ICDE}
\label{sec:org4fa01c4}
Scalability often overlooked but is a major cause of application failure.
Also hard to predict/test/validate.

Reliance on proven designs and technologies is essential.

For ICDE, application should be capable of handling a peak load of 150 concurrent requests from ICDE
clients.
Relatively easy to simulate user load to validate this.
\subsection{Modifiability}
\label{sec:org3124182}
Modifications to a software system during its lifetime are certain.
Modifiable systems are easier to change/evolve.

Assess modifiability in context of how a system is likely to change.
No need to facilitate changes that are unlikely to occur, so avoid over-engineering.

Modifiability measures how easy it may be to change an application to cater for new functional
or non-functional requirements.

Must estimate cost and effort of modifiability.

Scenarios could involve new features, replacing components, or porting applications.
\subsubsection{Analysis}
\label{sec:org8b5acd3}
Impact not easy to quantify.

Best possible is:
\begin{itemize}
\item convincing impact analysis of changes needed
\item demo of how solution can accommodate modification without change
\end{itemize}

Minimizing dependencies increases modifiablity.
Changes isolated to single components are likely to be less expensive than those that affect other
components as well.
\subsubsection{ICDE}
\label{sec:org27d4958}
Range of events trapped and stored by ICDE client to be expanded.

3rd party tools communicate new message types.

Change DB or server technology user.
\subsection{Security}
\label{sec:org557f510}
Difficult and specialized multi-faceted quality attribute.

Much technology available, but requires deep knowledge of approaches and solution.

\textbf{Authentication}: applications can verify identity of users and other applications with which
they communicate

\textbf{Authorization}: authenticated users and applications have defined access rights to system resources

\textbf{Encryption}: messages sent to/from application are encrypted

\textbf{Integrity}: ensures message contents are not altered in transit

\textbf{Non-repudiation}: message sender has proof of delivery and receiver is assured of sender's identity,
so neither can subsequently refute participation in message exchange

Approaches include:
\begin{itemize}
\item internet application security (SSL, PKI)
\item authentication and authorization in Java (JAAS)
\item OS security
\item DB security
\end{itemize}
\subsubsection{ICDE Requirements}
\label{sec:orga12c1bb}
Authentication of ICDE users and 3rd party ICDE tools to ICDE server.

Encryption of data to ICDE server from 3rd party tools/users executing remotely over
an insecure network.
\subsection{Availability}
\label{sec:orged32bd6}
Requirement for most IT applications.

Measured by the proportion of required time it is usable.

Related to reliability, unreliable \(\Rightarrow\) poor availability.

Period of loss of availability determined by:
\begin{itemize}
\item time to detect failure
\item time to correct failure
\item time to restart application
\end{itemize}

Strategies for high availability:
\begin{itemize}
\item eliminate single points of failure
\item replication and failover
\item automatic detection and restart
\end{itemize}

\textbf{Recoverability}: capability to reestablish performance levels and recover affected data after an
application or system failure
\subsubsection{ICDE}
\label{sec:org25a4109}
Achieve 100\% availability during business hours.

Plenty of scope for downtime for system upgrade, backup, and maintenance.

Include mechanisms for component replication and failover.
\subsection{Integration}
\label{sec:orgc3be8ca}
Ease with which application can be incorporated into broader application context.

Achieved by programmatic APIs and data integration.
\subsubsection{Strategies}
\label{sec:org042e6da}
Expose application data for access by other components.

Offer services to read/write application data through an abstracted interface.
\subsubsection{ICDE}
\label{sec:org55d3646}
Revolve around need to support 3rd party analysis tools.

Well-defined and understood mechanism for 3rd party tools to access data in the ICDE data store.
\subsection{Other Quality Attributes}
\label{sec:org8c5c568}
\subsubsection{Portability}
\label{sec:org2973397}
If an application can be easily executed on a different software/hardware platform to the one
it has been developed for.
\subsubsection{Testability}
\label{sec:org1503e0e}
The ease at which an application can be tested.
\subsubsection{Supportablity}
\label{sec:orgcb7ae82}
The ease at which an application can be supported once it is deployed.
\section{Design Trade-offs}
\label{sec:orga72750b}
Quality attributes interact with and affect each other.

Architects must create solutions that make sensible compromises since it is not possible to satisfy
all competing requirements.

Still, must satisfy all stakeholder needs, which is difficult.
\end{document}
